---
title: "Integration and Downstream analysis, remove hashtag-doublet, all samples"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: no
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: true
    theme: lumen
---


```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=14, fig.height=10}
#####----------------------------------------------------------------------#####
#
# trnguyen@ukaachen.de
#
#####----------------------------------------------------------------------#####

##### clean up #####
# gc()
# rm(list = ls())

scrna_pipeline_src <- "/home/hieunguyen/CRC1382/src_2023/src_pipeline/scRNA_GEX_pipeline/processes_src"
source(file.path(scrna_pipeline_src, "import_libraries.R"))
source(file.path(scrna_pipeline_src, "helper_functions.R"))

#####----------------------------------------------------------------------#####
# CONFIGURATIONS 
#####----------------------------------------------------------------------#####
analysis.round <- "1st"

chosen.seed <- 42
num.dim.integration <- 25 
num.PCA <- 25
num.dim.cluster <- 25
num.PC.used.in.Clustering <- 25

path.to.src <- "/home/hieunguyen/CRC1382/src/src_pipeline/scRNA_GEX_pipeline/processes_src"

# outdir <- "/media/hieunguyen/CRC1382H/CRC1382/outdir"
outdir <- "/home/hieunguyen/CRC1382/outdir"
# PROJECT <- "EStange_20240411_SeuratV4"
PROJECT <- "EStange_20240411_SeuratV4"
config.version <- "reduced_RNAcontam_0"
PROJECT <- sprintf("%s_%s", PROJECT, config.version)

path.to.main.output <- file.path(outdir, PROJECT, "data_analysis")

path.to.01.output <- file.path(path.to.main.output, "01_output")
path.to.02.output <- file.path(path.to.main.output, "02_output")
path.to.03.output <- file.path(path.to.main.output, "03_output")
dir.create(path.to.03.output, showWarnings = FALSE, recursive = TRUE)

all.samples <- c("adult",
                 "adult_GF",
                 "adult_SPF",
                 "d10_SPF",
                 "d15_SPF",
                 "d20_SPF",
                 "d4_GF",
                 "d4_LPS",
                 "d4_SPF",
                 "d7",
                 "d7_GF",
                 "d7_SPF")

all.2nd.run.samples <- setdiff(all.samples, c("adult", "d7"))
cluster.resolution <- 0.5
```



# Data integration

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if (file.exists(file.path(path.to.03.output, "all_sobj.rds")) == FALSE){
  all.s.obj <- list()

  for (sample.id in all.2nd.run.samples){
    print(sprintf("Working on sample %s", sample.id))
    all.s.obj[[sample.id]]<- readRDS(file.path(path.to.02.output, sample.id, sprintf("GEX_sample_%s_seurat_object.rds", sample.id)))
  }
  

  s.obj <- merge(x = all.s.obj[[1]], 
                   y = unlist(all.s.obj[2:length(all.s.obj)]),
                   merge.data = FALSE, 
                   add.cell.ids = names(all.s.obj))

  chosen.assay <- "RNA"
  DefaultAssay(s.obj) <- chosen.assay
  
  s.obj <- NormalizeData(s.obj) # ---> use Log Normalized
  s.obj <- FindVariableFeatures(s.obj, selection.method = "vst")
  s.obj <- ScaleData(s.obj, features = rownames(s.obj))
  
  s.obj <- RunPCA(s.obj, npcs = num.PCA, verbose = FALSE, reduction.name=sprintf("%s_PCA", chosen.assay))
  s.obj <- RunUMAP(s.obj, reduction = sprintf("%s_PCA", chosen.assay), dims = 1:num.PCA, 
                             reduction.name=sprintf("%s_UMAP", chosen.assay), seed.use = chosen.seed)
  s.obj <- RunTSNE(s.obj, reduction = sprintf("%s_PCA", chosen.assay), dims = 1:num.PCA, 
                             reduction.name=sprintf("%s_TSNE", chosen.assay), seed.use = chosen.seed)
  
  s.obj <- FindNeighbors(s.obj, reduction = sprintf("%s_PCA", chosen.assay), dims = 1:num.PC.used.in.Clustering)
  
  s.obj <- FindClusters(s.obj, resolution = cluster.resolution, random.seed = chosen.seed)
  
  saveRDS(object = s.obj, file = file.path(path.to.03.output, "all_sobj.rds"))  
  print("Finished saving the file")
  write.csv(data.frame(data = c("finished saving the object to disk")), file.path(path.to.03.output, "CHECK_all_sobj.rds.csv"))
} else {
  s.obj <- readRDS(file.path(path.to.03.output, "all_sobj.rds"))
}
```


# UMAP of all samples {.tabset}

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (sample.id in unique(s.obj$name)){
  cat(sprintf("## Sample: %s \n", sample.id))
  p <- DimPlot(object = subset(s.obj, name == sample.id), reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, label.size = 10, repel = TRUE) + 
  ggtitle(sprintf("UMAP Sample %s: All clusters", sample.id)) +
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
  print(p)
  cat("\n \n")
}

```

# Before and After integration

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
if (file.exists(file.path(path.to.03.output, "all_sobj.integrated.rds")) == FALSE){
  data.list <- SplitObject(s.obj, split.by = "name")
  data.list <- lapply(X = data.list, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)})
  
  k.filter <- 200
  
  anchors <- FindIntegrationAnchors(object.list = data.list, dims = 1:num.dim.integration, scale=F,
                                    k.filter = k.filter) ## THIS IS CCA DIMENSIONS
  
  s.obj_inte <- IntegrateData(anchorset = anchors, dims = 1:num.dim.integration, k.weight = k.filter) ## THIS IS PCA DIMENSION
  
  ## keep the order of integration obj
  s.obj_inte <- s.obj_inte[, colnames(s.obj)]
  
  s.obj[['integrated']] <- s.obj_inte[['integrated']]
  
  s.obj@commands <- c(s.obj@commands, s.obj_inte@commands)
  
  s.obj@tools <- c(s.obj@tools, s.obj_inte@tools)
  
  DefaultAssay(s.obj) <- "integrated"
  
  s.obj <- ScaleData(s.obj, verbose = FALSE, features = row.names(s.obj))
  
  s.obj <- RunPCA(s.obj, npcs = num.PCA, verbose = FALSE, reduction.name="INTE_PCA")
  
  s.obj <- RunUMAP(s.obj, reduction = "INTE_PCA", dims = 1:num.PCA, reduction.name="INTE_UMAP")
  
  # clustering 
  s.obj <- FindNeighbors(s.obj, reduction = "INTE_PCA", dims = 1:num.dim.cluster)
  
  s.obj <- FindClusters(s.obj, resolution = cluster.resolution)
  
  saveRDS(object = s.obj, file = file.path(path.to.03.output, "all_sobj.integrated.rds"))
  
  print("Finished saving the file")
  write.csv(data.frame(data = c("finished saving the object to disk")), file.path(path.to.03.output, "CHECK_all_sobj.integrated.rds.csv"))
} else {
  s.obj <- readRDS(file.path(path.to.03.output, "all_sobj.integrated.rds"))
}
```

## Before integration

## Integrate all samples and show them on UMAP

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, group.by = "name", label.box = TRUE, label.size = 10, repel = TRUE) +
  ggtitle("UMAP: all samples before performing data integration") + 
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, group.by = "name", label.box = TRUE, label.size = 10, repel = TRUE) +
  ggtitle("UMAP: all samples after performing data integration") + 
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = FALSE, split.by = "name", ncol = 3, label.box = TRUE, label.size = 10, repel = TRUE) +
  ggtitle("UMAP: all samples after performing data integration") + 
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```

## Clustering after integration
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
DimPlot(object = s.obj, reduction = "INTE_UMAP", label = TRUE, label.box = TRUE, label.size = 10, repel = TRUE) +
  ggtitle("UMAP: all samples after performing data integration") + 
  theme(axis.text = element_text(size=20, face = "bold"),
        axis.title = element_text(size=20, face = "bold"), 
        title = element_text(size = 20, face = "bold"))
```

# Ambient RNA contamination on UMAP 

## All samples, integrated
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
FeaturePlot(object = s.obj, features = "AmbientRNA", label = TRUE, label.size = 8, reduction = "INTE_UMAP")
```

## Each sample, integrated UMAP {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (sample.id in unique(s.obj$name)){
  cat(sprintf("### Sample %s \n", sample.id))
  p <- FeaturePlot(object = subset(s.obj, name == sample.id), features = "AmbientRNA", label = TRUE, label.size = 8, reduction = "INTE_UMAP")
  print(p)
  cat("\n \n")
}
```

## Each sample, Violin plot ambient RNA contamination level in each cluster {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}
for (sample.id in unique(s.obj$name)){
  cat(sprintf("### Sample %s \n", sample.id))
  p <- VlnPlot(object = subset(s.obj, name == sample.id), features = "AmbientRNA")
  print(p)
  cat("\n \n")
}
```

# Cluster markers of integrated dataset
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}

if (file.exists(file.path(path.to.03.output, "integrated_cluster_markers.rds")) == FALSE){
  integrated.cluster.markers <- FindAllMarkers(object = s.obj, assay = "RNA", test.use = "wilcox")
  saveRDS(integrated.cluster.markers, file.path(path.to.03.output, "integrated_cluster_markers.rds"))   
} else {
  integrated.cluster.markers <- readRDS(file.path(path.to.03.output, "integrated_cluster_markers.rds"))
}

integrated.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$p_val_adj <= 0.05 & integrated.cluster.markers$avg_log2FC > 0)
```

## Top-12 cluster marker genes `FeaturePlot` {.tabset}
Here we just show the top-12 genes (ranked by `avg_log2FC`) for each cluster. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- FeaturePlot(object = s.obj, reduction = "INTE_UMAP", features = head(tmp.cluster.markers, 9)$gene, ncol = 3, label = TRUE, pt.size = 0.5, label.size = 5, label.color = "red")  
  print(p)
  cat("\n \n")
}
```


## Top-12 dot plot {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- DotPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene)  
  print(p)
  cat("\n \n")
}
```

## Top-12 Violin plot {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=28, fig.height=20}
for (cluster.id in sort(unique(s.obj@meta.data$seurat_clusters))){
  DefaultAssay(s.obj) <- "RNA"
  cat(sprintf("### Cluster %s \n", cluster.id))
  tmp.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  p <- VlnPlot(object = s.obj, features = head(tmp.cluster.markers, 9)$gene)  
  print(p)
  cat("\n \n")
}
```

## Tables of all cluster marker genes {.tabset}
```{r echo=FALSE, include=FALSE}
integrated.cluster.markers %>% create_dt()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=14, fig.height=10}

for (cluster.id in sort(unique(integrated.cluster.markers$cluster))){
  tmp.integrated.cluster.markers <- subset(integrated.cluster.markers, integrated.cluster.markers$cluster == cluster.id) %>% arrange(desc(avg_log2FC))
  tmp.integrated.cluster.markers <- tmp.integrated.cluster.markers %>% mutate_if(is.numeric, round, 6) 
  
  cat(paste("\n\n### Cluster: ", cluster.id, "##\n"))
  print( htmltools::tagList(datatable(tmp.integrated.cluster.markers, extensions = 'Buttons',
                filter = "top",
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All")),
                               columnDefs = list(list(
                                 targets = "_all",
                                 render = JS(
                                   "function(data, type, row, meta) {",
                                   "return type === 'display' && data != null && data.length > 100 ?",
                                   "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                   "}")
                               ))
                ))))
  cat("\n \n")
}

```

# Number of cells in each cluster
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
count.cell.in.cluster <- s.obj@meta.data %>% subset(select = c(name, seurat_clusters)) %>% table() %>% as.data.frame()
count.cell.in.cluster.wider <- count.cell.in.cluster %>% pivot_wider(names_from = "name", values_from = "Freq") 

pct.table <- count.cell.in.cluster.wider %>% column_to_rownames("seurat_clusters")
colsum.table <- colSums(pct.table)
rowsum.table <- rowSums(pct.table)
names(rowsum.table) <- to_vec(for (item in names(rowsum.table)) sprintf("cluster_%s", item))

pct.table.sample <- pct.table # distribution of cells from each cluster in each sample
pct.table.cluster <- pct.table %>% t() %>% as.data.frame() # distribution of cell from each sample in each cluster
colnames(pct.table.cluster) <- to_vec(for (item in colnames(pct.table.cluster)) sprintf("cluster_%s", item))

for (sample.id in colnames(pct.table.sample)){
  pct.table.sample[[sample.id]] <- to_vec( for(item in pct.table.sample[[sample.id]]) item/colsum.table[[sample.id]])
}

for (cluster.id in colnames(pct.table.cluster)){
  pct.table.cluster[[cluster.id]] <- to_vec( for(item in pct.table.cluster[[cluster.id]]) item/rowsum.table[[cluster.id]])
}
```

## Table
### Raw counts 
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
count.cell.in.cluster.wider %>% create_dt()
```

### Percentage in each cluster
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.cluster %>% create_dt()
```

### Percentage in each sample
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.cluster %>% create_dt()
```

## Barplot

### Raw counts
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
p <- count.cell.in.cluster %>% ggplot(aes(x = seurat_clusters, y = Freq, fill = name)) + geom_bar(stat = "identity", dodge = "stack")
ggplotly(p)
```

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
p <- count.cell.in.cluster %>% ggplot(aes(x = name, y = Freq, fill = seurat_clusters)) + geom_bar(stat = "identity", dodge = "stack")
ggplotly(p)
```

### Percentages

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.sample.longer <- pct.table.sample %>% rownames_to_column("seurat_clusters") %>% 
  pivot_longer(!seurat_clusters, names_to = "name", values_to = "Freq")
pct.table.sample.longer$seurat_clusters <- factor(pct.table.sample.longer$seurat_clusters, levels = seq(0, length(unique(pct.table.sample.longer$seurat_clusters)) - 1))
p <- pct.table.sample.longer  %>% 
  ggplot(aes(x = name, y = Freq, fill = seurat_clusters)) + geom_bar(stat = "identity", dodge = "stack")
ggplotly(p)
```


```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.cluster.longer <- pct.table.cluster %>% rownames_to_column("name") %>% 
  pivot_longer(!name, names_to = "seurat_clusters", values_to = "Freq")

p <- pct.table.cluster.longer  %>% 
  ggplot(aes(x = seurat_clusters, y = Freq, fill = name)) + geom_bar(stat = "identity", dodge = "stack")
ggplotly(p)
```


# Number of hashtag in each cluster
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
count.cell.in.cluster <- s.obj@meta.data %>% subset(select = c(HTO_classification, seurat_clusters)) %>% table() %>% as.data.frame()
count.cell.in.cluster.wider <- count.cell.in.cluster %>% pivot_wider(names_from = "HTO_classification", values_from = "Freq") 

pct.table <- count.cell.in.cluster.wider %>% column_to_rownames("seurat_clusters")
colsum.table <- colSums(pct.table)
rowsum.table <- rowSums(pct.table)
names(rowsum.table) <- to_vec(for (item in names(rowsum.table)) sprintf("cluster_%s", item))

pct.table.sample <- pct.table # distribution of cells from each cluster in each sample
pct.table.cluster <- pct.table %>% t() %>% as.data.frame() # distribution of cell from each sample in each cluster
colnames(pct.table.cluster) <- to_vec(for (item in colnames(pct.table.cluster)) sprintf("cluster_%s", item))

for (sample.id in colnames(pct.table.sample)){
  pct.table.sample[[sample.id]] <- to_vec( for(item in pct.table.sample[[sample.id]]) item/colsum.table[[sample.id]])
}

for (cluster.id in colnames(pct.table.cluster)){
  pct.table.cluster[[cluster.id]] <- to_vec( for(item in pct.table.cluster[[cluster.id]]) item/rowsum.table[[cluster.id]])
}
```

## Table
### Raw counts 
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
count.cell.in.cluster.wider %>% create_dt()
```

### Percentage in each cluster
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.cluster %>% create_dt()
```

### Percentage in each sample
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.cluster %>% create_dt()
```

## Barplot

### Raw counts
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
p <- count.cell.in.cluster %>% ggplot(aes(x = seurat_clusters, y = Freq, fill = HTO_classification)) + geom_bar(stat = "identity", dodge = "stack") + theme(axis.text.x = element_text(angle = 90))
ggplotly(p)
```

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
p <- count.cell.in.cluster %>% ggplot(aes(x = HTO_classification, y = Freq, fill = seurat_clusters)) + geom_bar(stat = "identity", dodge = "stack") + theme(axis.text.x = element_text(angle = 90))
ggplotly(p)
```

### Percentages

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.sample.longer <- pct.table.sample %>% rownames_to_column("seurat_clusters") %>% 
  pivot_longer(!seurat_clusters, names_to = "name", values_to = "Freq")
pct.table.sample.longer$seurat_clusters <- factor(pct.table.sample.longer$seurat_clusters, levels = seq(0, length(unique(pct.table.sample.longer$seurat_clusters)) - 1))
p <- pct.table.sample.longer  %>% 
  ggplot(aes(x = name, y = Freq, fill = seurat_clusters)) + geom_bar(stat = "identity", dodge = "stack") + theme(axis.text.x = element_text(angle = 90))
ggplotly(p)
```


```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.cluster.longer <- pct.table.cluster %>% rownames_to_column("name") %>% 
  pivot_longer(!name, names_to = "seurat_clusters", values_to = "Freq")

p <- pct.table.cluster.longer  %>% 
  ggplot(aes(x = seurat_clusters, y = Freq, fill = name)) + geom_bar(stat = "identity", dodge = "stack") + theme(axis.text.x = element_text(angle = 90))
ggplotly(p)
```

# Number of sample-hashtag in each cluster
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
meta.data <- s.obj@meta.data %>% subset(select = c(seurat_clusters, name, HTO_classification)) %>%
  rowwise() %>%
  mutate(name_ht = sprintf("%s_%s", name, str_split(HTO_classification, "-")[[1]][[1]]))

count.cell.in.cluster <- meta.data %>% subset(select = c(name_ht, seurat_clusters)) %>% table() %>% as.data.frame()
count.cell.in.cluster.wider <- count.cell.in.cluster %>% pivot_wider(names_from = "name_ht", values_from = "Freq") 

pct.table <- count.cell.in.cluster.wider %>% column_to_rownames("seurat_clusters")
colsum.table <- colSums(pct.table)
rowsum.table <- rowSums(pct.table)
names(rowsum.table) <- to_vec(for (item in names(rowsum.table)) sprintf("cluster_%s", item))

pct.table.sample <- pct.table # distribution of cells from each cluster in each sample
pct.table.cluster <- pct.table %>% t() %>% as.data.frame() # distribution of cell from each sample in each cluster
colnames(pct.table.cluster) <- to_vec(for (item in colnames(pct.table.cluster)) sprintf("cluster_%s", item))

for (sample.id in colnames(pct.table.sample)){
  pct.table.sample[[sample.id]] <- to_vec( for(item in pct.table.sample[[sample.id]]) item/colsum.table[[sample.id]])
}

for (cluster.id in colnames(pct.table.cluster)){
  pct.table.cluster[[cluster.id]] <- to_vec( for(item in pct.table.cluster[[cluster.id]]) item/rowsum.table[[cluster.id]])
}
```

## Table
### Raw counts 
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
count.cell.in.cluster.wider %>% create_dt()
```

### Percentage in each cluster
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.cluster %>% create_dt()
```

### Percentage in each sample
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.cluster %>% create_dt()
```

## Barplot

### Raw counts
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
p <- count.cell.in.cluster %>% ggplot(aes(x = seurat_clusters, y = Freq, fill = name_ht)) + geom_bar(stat = "identity", dodge = "stack")
ggplotly(p)
```

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
p <- count.cell.in.cluster %>% ggplot(aes(x = name_ht, y = Freq, fill = seurat_clusters)) + geom_bar(stat = "identity", dodge = "stack") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggplotly(p)
```

### Percentages

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.sample.longer <- pct.table.sample %>% rownames_to_column("seurat_clusters") %>% 
  pivot_longer(!seurat_clusters, names_to = "name", values_to = "Freq")
pct.table.sample.longer$seurat_clusters <- factor(pct.table.sample.longer$seurat_clusters, levels = seq(0, length(unique(pct.table.sample.longer$seurat_clusters)) - 1))
p <- pct.table.sample.longer  %>% 
  ggplot(aes(x = name, y = Freq, fill = seurat_clusters)) + geom_bar(stat = "identity", dodge = "stack") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggplotly(p)
```

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=10}
pct.table.cluster.longer <- pct.table.cluster %>% rownames_to_column("name") %>% 
  pivot_longer(!name, names_to = "seurat_clusters", values_to = "Freq")

p <- pct.table.cluster.longer  %>% 
  ggplot(aes(x = seurat_clusters, y = Freq, fill = name)) + geom_bar(stat = "identity", dodge = "stack") +
  theme(axis.text.x = element_text(angle = 90))
ggplotly(p)
```

